---
title: "a3_task3_caitie_reza"
author: "Caitie Reza"
date: "11/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(ggplot2)
```

### Summary

####    In this document, I will explore, visualize, and statistically analyze purple sea urchin data (urchins.csv) from the Santa Barbara Coastal LTER site. The data was taken as part of the SBC LTER Reef Long-term experiment: Kelp removal: Urchin size frequency distribution.  See below for citation: 

Santa Barbara Coastal LTER and D. Reed. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b (Accessed 2020-11-03).

```{r}
urchin_data <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```


### Section 1. Comparing Urchin Size
##### In this section, I will explore the sizes of purple and red sea urchins by compare their distributions using several graph types (histogram, boxplot). Then, I will create a table with various summary statistics of the urchins data (read in as "urchin_data").

#### 1a. Visualization

```{r}
ggplot(data = urchin_data, aes(x = size, fill = common_name)) +
  geom_histogram() +
  facet_wrap(~ common_name) +
  theme(legend.position = "none") +
  labs(x = "Size (cm)", y = "Count (Individual Urchins)",
       title = "Fig. 1 Size Distribution of Purple and Red Urchins at SBC LTER Site")
```


```{r}
ggplot(data = urchin_data, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Common Name", y = "Size (cm)",
       title = "Fig. 2 Average Size of Purple and Red Urchins at SBC LTER Site") +
  theme_gray()

```

#### 1b. Summary Statistics
Here, I will create a summary table that contains the mean, median, standard deviation, and sample size for purple and red urchins.


```{r}
urchin_summary <- urchin_data %>% 
  group_by(common_name) %>% 
  summarize(
    av_size = mean(size),
    median_size = median(size),
    standard_dev = sd(size),
    sample_size = n())

urchin_summary

```


#### Figures 1 and 2, Summary Stats Table

The mean red urchin size is 6.77 +/- 1.89 cm while that of purple urchins is 4.47 +/- 1.18 cm, resulting in a 2.3 cm  difference in average size between the species. The purple urchin distribution (larger sample size) is negatively skewed, while the distribution of red urchin size (smaller sample) is more normal.  








### Section 2. Purple Urchin Counts by Site/Year

In this section, I will be exploring counts of purple urchins at each site by year. 


#### 2a. purple_urchin Subset
In a piped sequence, I will create a stored subset of the urchins data called purple_urchins where I: 
  1) Filter to only include observations for purple urchins
  2) Convert the date to class “Date” 
  3) Create a new column that only contains the year
  4) Create a new column called site_full that contains the full name (not abbreviation) for each of the five site

```{r}
purple_urchins <- urchin_data %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(Date = mdy(date)) %>% 
  mutate(Year = year(Date)) %>% 
  mutate(site_full = case_when(
     site %in% "NAPL" ~ "Naples Reef",
     site %in% "AQUE" ~ "Arroyo Quemado Reef",
     site %in% "CARP" ~ "Carpinteria Reef",
     site %in% "MOHK" ~ "Mohawk Reef",
     site %in% "IVEE" ~ "Isla Vista Reef"))
  

```


#### 2b. Purple Urchin Summary Table

Here, I will make a summary table with urchin counts for each site taken each year. 


```{r}
purple_summary <- purple_urchins %>% 
  group_by(Year, site_full) %>% 
  count(site_full) %>% 
  rename("Urchins" = "n")

purple_summary
  
```

#### 2c. Purple Urchin Graph

Here, I will make 5 separate graphs (one for each site) using ggplot to explore the urchin counts each year at every site, using the table I summarized above. I will use facet_wrap to separate by site. 



```{r}
ggplot(data = purple_summary, aes(x = Year, y = Urchins)) +
  geom_col(aes(fill = site_full)) +
  facet_wrap(~site_full) +
  theme(legend.position = "none") +
  labs(y = "Individual Urchins",
       title = "Purple Urchin Counts at 5 Sites in the SBC LTER")
```







